Pre-/Post-Trikafta in CFRD

### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
pre post ETI Chan qmd instead of rmd

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gtsummary)
library(arsenal)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Baseline HbA1c, BMI group pre- and post-modulator
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
  ) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
# Pre-modulator BMI percentile
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
# Columns formats
df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
```

>>>>

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(gtsummary)
library(arsenal)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# CHCO data
chco_bmi_pft <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv",
  na.strings = ""
)
chco_bmi_pft$Date <- mdy(chco_bmi_pft$Date)
chco_cgm <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/chco_cgm.csv",
  na.strings = ""
)
chco_cgm$Date <- date(parse_date_time(chco_cgm$date_cgm_placement,
  orders = c("mdyHM", "mdy")
))
chco_glycemic <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv",
  na.strings = ""
)
chco_glycemic$Date <- mdy(chco_glycemic$Date)
chco_registry <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/chco_registry.csv",
  na.strings = ""
)
# Montana
montana_bmi_pft <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv",
  na.strings = ""
)
montana_cgm <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_cgm.csv",
  na.strings = ""
)
montana_cgm$Date <- date(parse_date_time(montana_cgm$date_cgm_placement,
  orders = c("mdyHM", "mdy")
))
montana_demo <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_demo.csv",
  na.strings = ""
)
montana_genotypes <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv",
  na.strings = ""
)
montana_gtubes <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv",
  na.strings = ""
)
montana_pancreatic <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv",
  na.strings = ""
)
montana_glycemic_registry <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv",
  na.strings = ""
)
montana_glycemic_registry$Date <- mdy(montana_glycemic_registry$Date)
# UW
uw_registry <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
  na.strings = ""
)
uw_bmi <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_bmi.csv",
  na.strings = ""
)
uw_pft <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_pft.csv",
  na.strings = ""
)
uw_cgm <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_cgm.csv",
  na.strings = ""
)
uw_demo <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_demo.csv",
  na.strings = ""
)
uw_genotypes <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
  na.strings = ""
)
uw_glycemic <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv",
  na.strings = ""
)
uw_gtubes <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv",
  na.strings = ""
)
uw_pancreatic <- read.csv(
  "./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv",
  na.strings = ""
)
```

```{r}
# Baseline HbA1c, BMI group pre- and post-modulator
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
  ) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
# Pre-modulator BMI percentile
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
# Columns formats
df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
```
